### 2014-11-24 manuscript/AppendixA.tex
Added ms text on simulation; now fully specify Gibbs samplers in Appendix A; started appendix B for simulation study details.
Some modifications to code for simulating data, including plotting code

\def\be{{\ensuremath\mathbf{e}}}
\def\bx{{\ensuremath\mathbf{x}}}
\def\bX{{\ensuremath\mathbf{X}}}
\def\bthet{{\ensuremath\boldsymbol{\theta}}}
\newcommand{\VS}{V\&S}
\newcommand{\tr}{\mbox{tr}}
%\renewcommand{\refname}{\hspace{2.3in} \normalfont \normalsize LITERATURE CITED}
%this tells it to put 'Literature Cited' instead of 'References'
\bibpunct{(}{)}{,}{a}{}{;}
\oddsidemargin 0.0in
\evensidemargin 0.0in
\textwidth 6.5in
\headheight 0.0in
\topmargin 0.0in
\textheight=9.0in
%\renewcommand{\tablename}{\textbf{Table}}
%\renewcommand{\figurename}{\textbf{Figure}}
\renewcommand{\em}{\it}
\renewcommand\thefigure{S1.\arabic{figure}}

@@

\begin{center}
Appendix S1: Model formulation and Gibbs sampling algorithms for certain classes and extensions of the generalized linear model
\bigskip
\end{center}
\vspace{.3in}

@@

We write all statistical models for ecological prediction in the form
\begin{linenomath*}
\begin{equation}
  \label{eq:basic_model}
  Y_i \sim f_Y(g^{-1}(\mu_i)),
\end{equation}
\end{linenomath*}
where $\mu_i=\theta_i + \epsilon_i$, $f_Y$ denotes a probability density or mass function (e.g. Bernoulli, Poisson), $g$ gives a link function,
$\theta_i$ is a linear predictor, and $\epsilon_i \sim {\rm Normal}(0,\tau_\epsilon)$ is iid Gaussian error with precision parameter $\tau_\epsilon$.  The specification in Eq. \ref{eq:basic_model} is thus doubly stochastic, in the sense that we assume error associated with $f_Y$ as well as in the location parameter $\mu_i$.  This setup can be useful computationally, and can also be used to approximate singly stochastic system by setting $\tau_\epsilon$ to a large value.  For instance, models for count data often assume a Poisson error structure with a log link function to ensure that the Poisson intensity parameter is greater than zero.  In this case, we would specify
\begin{linenomath*}
\begin{equation*}
  Y_i \sim \mathrm{Poisson}(\exp(\theta_i + \epsilon_i)),
\end{equation*}
\end{linenomath*}
a configuration known as a log-Gaussian Cox process.
We now describe how different classes of statistical models can be developed depending on how one structures the linear predictor, $\theta_i$.

@@

Generalized linear models \citep{McCullaghNelder1989} are one of the simplest (and most often used) statistical models
used by ecologists to make spatial predictions. In GLMs, the linear predictor is a simple linear function of gathered covariates
(including possible quadratic terms of these covariates.  Statisticians often describe this relationship by
$\boldsymbol{\theta}={\bf X} \boldsymbol{\beta}$, which is written in matrix notation.  In particular, $\boldsymbol{\theta}$ gives
a vector of linear predictor values (one for each data point being analyzed), $\boldsymbol{\beta}$ gives a vector of regression coefficients, and ${\bf X}$ is a design matrix which includes all explanatory variables (and often a column vector of ones to represent an intercept).

@@

There are a number of ways smooth relationships can be modeled; in order to formulate a GAM in the notation of Eq. \ref{eq:basic_model} we employ a knot-based kernel smoother with a radial basis function.  The basic notion is to place $k_j$ knots, which we will denote by $\boldsymbol{\omega}_j = \{ \omega_{1j},\omega_{2j},\hdots,\omega_{k_j j} \}$, throughout the range of a given regressor $j$, where the number of knots controls the level of smoothing.  For each regressor of interest, we can then calculate a matrix ${\bf K}_j$ of dimension $(n,k_j)$, with entries $\mathcal{N}(X_{ij};\omega_j,\tau_{\omega,j})$, where $\mathcal{N}(x;\mu,\tau)$ denotes a Gaussian distribution with mean $\mu$ and precision $\tau$.  We can then introduce additional regression coefficients $\boldsymbol{\alpha}_j$ to help model the smooth relationship, and incorporate these into the linear predictor.  In the case of one smooth term, we then have  in Eq. \ref{eq:basic_model}:
\begin{linenomath*}
\begin{equation}
  \label{eq:GAM}
  \boldsymbol{\theta}={\bf X} \boldsymbol{\beta} + {\bf K}_j \boldsymbol{\alpha}_j
\end{equation}
\end{linenomath*}
When fitting such a model, one needs to select the number and location of knots, as well as the precision parameter $\tau_{\omega,j}$ for the Gaussian basis kernels.  This could be done using some notion of optimality or prediction error (as with cross validation).  In practice, a simple recommendation is to set $\tau_{\omega,j}$ less than the range of the modeled covariate, and greater than the typical difference between observed values of the covariate.

Our main point in writing the GAM as in Eq. \ref{eq:GAM} is to emphasize the structural similarity between GLMs and certain classes
of GAMs.  In particular,
we can rewrite Eq. \ref{eq:GAM} as
$\boldsymbol{\theta}={\bf X}_{aug} \boldsymbol{\beta}_{aug}$,
where ${\bf X}_{aug} = \left[ {\bf X} \hspace{3mm} {\bf K}_j \right]$ (i.e., concatenating ${\bf X}$ and ${\bf K}_j$ horizontally), and $\boldsymbol{\beta}_{aug}=\left[  \boldsymbol{\beta} \hspace{3mm} \boldsymbol{\alpha}_j \right] ^\prime$ (the subscript $aug$ denotes augmentation).  We shall exploit this structure when examining
extrapolation bias.  We note that similar formulations for GAMs (e.g. using thin plate splines) can be represented in a similar form \citep[cf.][]{CrainiceanuEtAl2005}.

@@

There are a variety of ways spatial autocorrelation can be included in regression models, depending on (i) the spatial support (i.e., continuous vs. discrete), and (ii) the particular mechanism used to impart correlation.  Here, we shall focus on areal models for discrete spatial support, as our impression is that these are more commonly employed in ecological studies.  Such models require that data are aggregated at the level of some sample unit (plots, grid cells, etc.).  Spatio-temporal regression models (STRMs) for areal data are often specified in a similar fashion to GLMs and GAMs:
\begin{linenomath*}
\begin{equation}
  \label{eq:spat_reg}
  \boldsymbol{\theta}={\bf X} \boldsymbol{\beta} + \boldsymbol{\eta},
\end{equation}
\end{linenomath*}
where $\boldsymbol{\eta}$ represent spatially autocorrelated effects.  In this treatment we shall consider two approaches for
inducing spatial autocorrelation in $\boldsymbol{\eta}$: process convolution
\citep[PC;][]{Higdon1998} and restricted spatial regression \citep[RSR;][]{Reich2006,Hodges2010,Hughes2013}.

The PC implementation works by placing $k$ knots throughout the spatial domain being analyzed, and like our GAM formulation, uses distances from the center point of each sample unit to each of these knot locations to induce spatial structure.  Using similar notation to our GAM formulation, we denote these knots as $\boldsymbol{\omega} = \{ \boldsymbol{\omega_{1}},\boldsymbol{\omega}_{2},\hdots,\boldsymbol{\omega}_{k} \}$; the $\boldsymbol{\omega}_m$ are bivariate in this  instance as space is assumed to be two dimensional.  We then construct a matrix ${\bf K}$ of dimension $(n,k)$ with entries $\mathcal{BVN}({\bf s}_i;\boldsymbol{\omega}_m,\tau_{\omega})$, where $\mathcal{BVN}({\bf x};{\bf \mu},\tau)$ denotes a bivariate Gaussian distribution with mean $\boldsymbol{\mu}$ and precision $\tau$.  As with our GAM model we introduce additional regression coefficients $\boldsymbol{\alpha}$ to estimate the level of spatial smoothing conditional on the assumed knot structure and reparameterize the STRM as
\begin{linenomath*}
\begin{equation}
  \label{eq:PC}
  \boldsymbol{\theta}={\bf X} \boldsymbol{\beta} + {\bf K} \boldsymbol{\alpha}
\end{equation}
\end{linenomath*}
This model is structurally very similar to the GAM model (indeed it may be interpreted as a GAM with a bivariate smooth on spatial location).  Similarly, we can rearrange terms to write it similar to the GLM, such that
$\boldsymbol{\theta}={\bf X}_{aug} \boldsymbol{\beta}_{aug}$,
where in this case ${\bf X}_{aug} = \left[ {\bf X} \hspace{3mm} {\bf K} \right]$, and $\boldsymbol{\beta}_{aug}=\left[  \boldsymbol{\beta} \hspace{3mm} \boldsymbol{\alpha} \right] ^\prime$.

@@

>>>>

\def\be{{\ensuremath\mathbf{e}}}
\def\bx{{\ensuremath\mathbf{x}}}
\def\bX{{\ensuremath\mathbf{X}}}
\def\bthet{{\ensuremath\boldsymbol{\theta}}}
\newcommand{\VS}{V\&S}
\newcommand{\tr}{\mbox{tr}}
%\renewcommand{\refname}{\hspace{2.3in} \normalfont \normalsize LITERATURE CITED}
%this tells it to put 'Literature Cited' instead of 'References'
\bibpunct{(}{)}{,}{a}{}{;}
\oddsidemargin 0.0in
\evensidemargin 0.0in
\textwidth 6.5in
\headheight 0.0in
\topmargin 0.0in
\textheight=9.0in
%\renewcommand{\tablename}{\textbf{Table}}
%\renewcommand{\figurename}{\textbf{Figure}}
\renewcommand{\em}{\it}
\renewcommand\thefigure{A\arabic{figure}}
\renewcommand\thetable{A\arabic{table}}
\renewcommand\theequation{A.\arabic{equation}}

@@

\begin{center}
Appendix A: Model formulation and Gibbs sampling algorithms for certain classes and extensions of the generalized linear model
\bigskip
\end{center}
\vspace{.3in}

@@

We write all statistical models for ecological prediction in the form
\begin{linenomath*}
\begin{equation}
  \label{eq:basic_model}
  Y_i \sim f_Y(g^{-1}(\mu_i)),
\end{equation}
\end{linenomath*}
where $\mu_i=\xi_i + \epsilon_i$, $f_Y$ denotes a probability density or mass function (e.g. Bernoulli, Poisson), $g$ gives a link function,
$\xi_i$ is a linear predictor, and $\epsilon_i \sim {\rm Normal}(0,\tau_\epsilon)$ is iid Gaussian error with precision parameter $\tau_\epsilon$.  The specification in Eq. \ref{eq:basic_model} is thus doubly stochastic, in the sense that we assume error associated with $f_Y$ as well as in the location parameter $\mu_i$.  This setup can be useful computationally, and can also be used to approximate singly stochastic system by setting $\tau_\epsilon$ to a large value.  For instance, models for count data often assume a Poisson error structure with a log link function to ensure that the Poisson intensity parameter is greater than zero.  In this case, we would specify
\begin{linenomath*}
\begin{equation*}
  Y_i \sim \mathrm{Poisson}(\exp(\xi_i + \epsilon_i)),
\end{equation*}
\end{linenomath*}
a configuration known as a log-Gaussian Cox process.
We now describe how different classes of statistical models can be developed depending on how one structures the linear predictor, $\xi_i$.

@@

Generalized linear models \citep{McCullaghNelder1989} are one of the simplest (and most often used) statistical models
used by ecologists to make spatial predictions. In GLMs, the linear predictor is a simple linear function of gathered covariates
(including possible polynomial terms and interactions).  Statisticians often describe this relationship by
$\boldsymbol{\xi}={\bf X} \boldsymbol{\beta}$, which is written in matrix notation.  In particular, $\boldsymbol{\xi}$ gives
a vector of linear predictor values (one for each data point being analyzed), $\boldsymbol{\beta}$ gives a vector of regression coefficients, and ${\bf X}$ is a design matrix which includes all explanatory variables (and often a column vector of ones to represent an intercept).

@@

\hspace{.5in}There are a number of ways smooth relationships can be modeled; in order to formulate a GAM in the notation of Eq. \ref{eq:basic_model} we employ a knot-based kernel smoother with a Gaussian radial basis function.  The basic notion is to place $k_j$ knots, which we will denote by $\boldsymbol{\omega}_j = \{ \omega_{1j},\omega_{2j},\hdots,\omega_{k_j j} \}$, throughout the range of a given regressor $j$, where the number of knots controls the level of smoothing.  For each regressor of interest, we can then calculate a matrix ${\bf K}_j$ of dimension $(n,k_j)$, with entries $\mathcal{N}(X_{ij};\omega_j,\tau_{\omega,j})$, where $\mathcal{N}(x;\mu,\tau)$ denotes a Gaussian distribution with mean $\mu$ and precision $\tau$.  We can then introduce additional regression coefficients $\boldsymbol{\alpha}_j$ to help model the smooth relationship, and incorporate these into the linear predictor.  In the case of one smooth term, we then have  in Eq. \ref{eq:basic_model}:
\begin{linenomath*}
\begin{equation}
  \label{eq:GAM}
  \boldsymbol{\xi}={\bf X} \boldsymbol{\beta} + {\bf K}_j \boldsymbol{\alpha}_j
\end{equation}
\end{linenomath*}
When fitting such a model, one needs to select the number and location of knots, as well as the precision parameter $\tau_{\omega,j}$ for the Gaussian basis kernels.  This could be done using some notion of optimality or prediction error (as with cross validation).  In practice, a simple recommendation is to set $\tau_{\omega,j}$ less than the range of the modeled covariate, and greater than the typical difference between observed values of the covariate.

\hspace{.5in}Our main point in writing the GAM as in Eq. \ref{eq:GAM} is to emphasize the structural similarity between GLMs and certain classes
of GAMs.  In particular,
we can rewrite Eq. \ref{eq:GAM} as
$\boldsymbol{\xi}={\bf X}_{aug} \boldsymbol{\beta}_{aug}$,
where ${\bf X}_{aug} = \left[ {\bf X} \hspace{3mm} {\bf K}_j \right]$ (i.e., concatenating ${\bf X}$ and ${\bf K}_j$ horizontally), and $\boldsymbol{\beta}_{aug}=\left[  \boldsymbol{\beta} \hspace{3mm} \boldsymbol{\alpha}_j \right] ^\prime$ (the subscript $aug$ denotes augmentation).  We shall exploit this structure when examining
extrapolation bias.  We note that similar formulations for GAMs (e.g. using thin plate splines) can be represented in a similar form \citep[cf.][]{CrainiceanuEtAl2005}.

@@

There are a variety of ways spatial autocorrelation can be included in regression models, depending on (i) the spatial support (i.e., continuous vs. discrete), and (ii) the particular mechanism used to impart correlation.  Here, we shall focus on areal models for discrete spatial support, as our impression is that these are more commonly employed in ecological studies.  Such models require that data are aggregated at the level of some sample unit (plots, grid cells, etc.).  Spatio-temporal regression models (STRMs) for areal data are often specified in a similar fashion to GLMs and GAMs:
\begin{linenomath*}
\begin{equation}
  \label{eq:spat_reg}
  \boldsymbol{\xi}={\bf X} \boldsymbol{\beta} + \boldsymbol{\eta},
\end{equation}
\end{linenomath*}
where $\boldsymbol{\eta}$ represent spatially autocorrelated effects.  In this treatment we shall consider two approaches for
inducing spatial autocorrelation in $\boldsymbol{\eta}$: process convolution
\citep[PC;][]{Higdon1998} and restricted spatial regression \citep[RSR;][]{Reich2006,Hodges2010,Hughes2013}.

The PC implementation works by placing $k$ knots throughout the spatial domain being analyzed, and like our GAM formulation, uses distances from the center point of each sample unit to each of these knot locations to induce spatial structure.  Using similar notation to our GAM formulation, we denote these knots as $\boldsymbol{\omega} = \{ \boldsymbol{\omega_{1}},\boldsymbol{\omega}_{2},\hdots,\boldsymbol{\omega}_{k} \}$; the $\boldsymbol{\omega}_m$ are bivariate in this  instance as space is assumed to be two dimensional.  We then construct a matrix ${\bf K}$ of dimension $(n,k)$ with entries $\mathcal{BVN}({\bf s}_i;\boldsymbol{\omega}_m,\tau_{\omega})$, where $\mathcal{BVN}({\bf x};{\bf \mu},\tau)$ denotes a bivariate Gaussian distribution with mean $\boldsymbol{\mu}$ and precision $\tau$.  As with our GAM model we introduce additional regression coefficients $\boldsymbol{\alpha}$ to estimate the level of spatial smoothing conditional on the assumed knot structure and reparameterize the STRM as
\begin{linenomath*}
\begin{equation}
  \label{eq:PC}
  \boldsymbol{\xi}={\bf X} \boldsymbol{\beta} + {\bf K} \boldsymbol{\alpha}
\end{equation}
\end{linenomath*}
This model is structurally very similar to the GAM model (indeed it may be interpreted as a GAM with a bivariate smooth on spatial location).  Similarly, we can rearrange terms to write it similar to the GLM, such that
$\boldsymbol{\xi}={\bf X}_{aug} \boldsymbol{\beta}_{aug}$,
where in this case ${\bf X}_{aug} = \left[ {\bf X} \hspace{3mm} {\bf K} \right]$, and $\boldsymbol{\beta}_{aug}=\left[  \boldsymbol{\beta} \hspace{3mm} \boldsymbol{\alpha} \right] ^\prime$.

@@

\section{Gibbs sampling}

We now describe methods for conducting Bayesian analysis for the models described above.  We utilize Gibbs sampling, characterized by cyclically sampling groups of parameters according to their so-called full conditional distributions conditionally on data and other parameters \citep{GelmanEtAl2004}.  In some cases, we are able to solve for full conditional distributions analytically and to sample from known families of probability distributions; in others, we resort to Metropolis-Hastings steps.

\subsection{Gibbs sampling for the GLM model}

Our Gibbs sampler for the GLM model consists of three sets of parameter updates (1) regression parameters ($\boldsymbol{\beta}$), (2) error precision ($\tau_\epsilon$), and (3) latent log-density ($\boldsymbol{\mu}$).

\underline{Sampling from $[\boldsymbol{\beta}|\cdot]$}

Setting the prior for the suite of regression parameters as $[\boldsymbol{\beta}] = \textrm{MVN}({\bf 0},(\tau_\beta X^\prime X)^{-1})$ results in the full conditional distribution
\begin{linenomath*}
\begin{equation*}
   [\boldsymbol{\beta} | \cdot] \equiv \mathcal{N}(({\bf X}^\prime {\bf X})^{-1}{\bf X}^\prime \boldsymbol{\mu},({\bf X}^\prime {\bf X})^{-1} (\tau_\epsilon + \tau_\beta)^{-1}).
\end{equation*}
\end{linenomath*}
We set $\tau_\beta = 0.01$ for applications described in this paper.

\underline{Sampling from $[\boldsymbol{\tau_\epsilon}|\cdot]$}

Specifying a conjugate $\text{Gamma}(a,b)$ prior for the precision parameter leads to a full conditional that is also Gamma distributed:
\begin{eqnarray}
  [\tau_\epsilon | \cdot] & \equiv & \textrm{Gamma}(0.5n + a,0.5 \boldsymbol{\Delta}^\prime \boldsymbol{\Delta} +b),
  \label{eq:tau_eps}
\end{eqnarray}
where $\Delta = \boldsymbol{\mu}-{\bf X}\boldsymbol{\beta}$.  For applications in this paper we set $a=1.0$ and $b=0.01$.  This selection admits a large range of permissable parameter values while maintaining flat probability mass near the origin.

\underline{Sampling from $[\boldsymbol{\mu}|\cdot]$}

Although $\boldsymbol{\mu}$ could potentially be integrated out of the likelihood, its inclusion allows for easier computation of full conditionals for $\boldsymbol{\beta}$ and $\tau_\epsilon$ (and for other full conditionals once the model is expanded, e.g. for GAMs and STRMs).  During the main MCMC phase, the $\boldsymbol{\mu}$ only need to be updated for grid cells that are exposed to sampling (see a subsequent section for information on posterior prediction for details on unsampled areas).  The full conditional for $\mu_s$ in a given grid cell $s$ is given by
\begin{equation*}
  [\mu_{s,t} | \cdot] \propto \mathcal{N}(\mu_s;{\bf X}_s \boldsymbol{\beta},\tau_\epsilon^{-0.5}) \times \textrm{Poisson}(C_s,\exp(o_s+\mu_s)),
\end{equation*}
where ${\bf X}_s$ gives the $s$th row of the design matrix and $o_s$ is an offset.  For our examples, quadrats are configured to represent $10\%$ of a grid cell, cell $o_s = \log(0.1)$.
We use a Metropolis-Hastings step to sample from $[\mu_{s,t} | \cdot]$.

\subsection{Gibbs sampling for the GAM model}

We used a Gibbs sampler for the GAM model that looks very familiar to that for the GLM model.

\underline{Sampling from $[\boldsymbol{\tau_\epsilon}|\cdot]$}

Assuming a Gamma($a,b$) prior, the full conditional for $\boldsymbol{\tau_\epsilon}$ is once again given by Eq. \ref{eq:tau_eps}.  In this case,
$\Delta = \boldsymbol{\mu}-{\bf X}\boldsymbol{\beta}-{\bf K}\boldsymbol{\alpha}$.

\underline{Sampling from $[\boldsymbol{\mu}|\cdot]$}

The full conditional for $\mu_s$ in a given grid cell $s$ is given by
\begin{equation*}
  [\mu_{s,t} | \cdot] \propto \mathcal{N}(\mu_s;{\bf X}\boldsymbol{\beta}+{\bf K}_s \boldsymbol{\alpha},\tau_\epsilon^{-0.5}) \times \textrm{Poisson}(C_s,\exp(o_s+\mu_s)),
\end{equation*}
where ${\bf K}_s$ gives the $s$th row of ${\bf K}$.

\underline{Sampling from $[\boldsymbol{\alpha}|\cdot]$}

We impose a Gaussian prior on smoothing kernel weights, $\boldsymbol{\alpha} \sim \mathcal{N}(0,\tau_\alpha^{-1})$.  Like $\boldsymbol{\beta}$, the resulting full conditional is once again in closed form:
\begin{linenomath*}
\begin{equation*}
   [\boldsymbol{\alpha} | \cdot] \equiv \mathcal{MVN}(({\bf K}^\prime {\bf K})^{-1}{\bf K}^\prime (\boldsymbol{\mu}-{\bf X}\boldsymbol{\beta}),({\bf K}^\prime {\bf K}^{-1} (\tau_\epsilon + \tau_\alpha)^{-1}).
\end{equation*}
\end{linenomath*}
We set $\tau_\alpha = 0.01$ in all subsequent applications.

\subsection{Gibbs sampler for the PC STRM model}

The Gibbs sampler for the process convolution version of the STRM model consists of the following:

\underline{Sampling from $[\boldsymbol{\beta}|\cdot]$}

Using the same prior as for the GLM results in the full conditional distribution
\begin{linenomath*}
\begin{equation*}
   [\boldsymbol{\beta} | \cdot] \equiv \mathcal{N}(({\bf X}^\prime {\bf X})^{-1}{\bf X}^\prime (\boldsymbol{\mu}-{\bf K}\boldsymbol{\alpha}),({\bf X}^\prime {\bf X})^{-1} (\tau_\epsilon + \tau_\beta)^{-1}).
\end{equation*}
\end{linenomath*}

\underline{Sampling from $[\boldsymbol{\mu}|\cdot]$}

The full conditional for $\mu_s$ in a given grid cell $s$ is given by
\begin{equation*}
  [\mu_{s,t} | \cdot] \propto \mathcal{N}(\mu_s;{\bf X}_s \boldsymbol{\beta}+{\bf K}_s \boldsymbol{\alpha},\tau_\epsilon^{-0.5}) \times \textrm{Poisson}(C_s,\exp(o_s+\mu_s)),
\end{equation*}
where ${\bf K}_s$ gives the $s$th row of ${\bf K}$.

\underline{Sampling from $[\tau_\epsilon|\cdot]$}

The full conditional is the same as for the GAM model with the replacement $\Delta = \boldsymbol{\mu}-{\bf X}\boldsymbol{\beta}-{\bf K}\boldsymbol{\alpha}$.

\underline{Sampling from $[\boldsymbol{\alpha}|\cdot]$}

We impose the same Gaussian prior on smoothing kernel weights, $\boldsymbol{\alpha} \sim \mathcal{N}(0,\tau_\alpha^{-1})$ as for the GAM.  The resulting full conditional is simply
\begin{linenomath*}
\begin{equation*}
   [\boldsymbol{\alpha} | \cdot] \equiv \mathcal{MVN}(({\bf K}^\prime {\bf K})^{-1}{\bf K}^\prime (\boldsymbol{\mu}-{\bf X}\boldsymbol{\beta}),({\bf K}^\prime {\bf K}^{-1} (\tau_\epsilon + \tau_\alpha)^{-1}).
\end{equation*}
\end{linenomath*}

\subsection{Gibbs sampler for the RSR STRM model}

Although ${\bf K}$ is defined differently in the RSR model, the Gibbs sampler remains the same as for the the PC model, with the exception of a modified full conditional for $[\boldsymbol{\alpha}|\cdot]$ owing to a different prior specification for $\boldsymbol{\alpha}$.  We must also update an additional precision parameter associated with spatial random effects, $\tau_\eta$.

\underline{Sampling from $[\boldsymbol{\alpha}|\cdot]$}

In the RSR model, the implied prior for reduced dimension spatial effects is $\boldsymbol{\alpha} \sim \text{MVN}(0,\tau_\eta^{-1}{\bf K}^\prime {\bf Q} {\bf K})$.  The resulting full conditional is
\begin{linenomath*}
\begin{eqnarray*}
   [\boldsymbol{\alpha} | \cdot] & = & \mathcal{MVN}({\bf M},\boldsymbol{\Sigma}), \text{ where} \\
   \boldsymbol{\Sigma}^{-1} & = & {\bf K}^\prime {\bf K}\tau_\epsilon+\tau_\eta{\bf K}^\prime {\bf Q} {\bf K} \text{ and} \\
    {\bf M} & = & \boldsymbol{\Sigma} \tau_\epsilon {\bf K}^\prime (\boldsymbol{\mu} - {\bf X} \boldsymbol{\beta}).
\end{eqnarray*}
\end{linenomath*}
See for example, \citet{ConnEtAl2014}.

\underline{Sampling from $\tau_\eta|\cdot]$}

Using a conjugate Gamma$(a,b)$ prior on $\tau_\eta$, we have
  \begin{equation*}
  \tau_{\eta} \sim {\rm Gamma}(0.5m + a,0.5 \boldsymbol{\alpha}^\prime
  {\bf K}^\prime {\bf Q} {\bf K}\boldsymbol{\alpha}+b).
  \end{equation*}
  Here, $m$ is the number of $\boldsymbol{\alpha}$ parameters.\\

\subsection{Generating posterior predictions}
If all we wanted to do was to estimate regression coefficients, the previous samplers would suffice.  However, we typically also wish to predict animal abundance across the landscape to come up with a total abundance estimate and produce a density map.  Fortunately, this is easy to do using posterior prediction, which consists of the following steps:
\begin{enumerate}
\item For each of $n$ samples (where $n$ is the desired posterior predictive sample size), choose an iteration of the MCMC sampler randomly with replacement (call this iteration $t$).
\item Label the $t$th sample of $\boldsymbol{\alpha}$, $\boldsymbol{\beta}$, $\boldsymbol{\tau_\epsilon}$, and $\boldsymbol{\mu}$ as $\boldsymbol{\alpha}^{(t)}$, $\boldsymbol{\beta}^{(t)}$, $\boldsymbol{\tau_\epsilon}^{(t)}$, and $\boldsymbol{\mu}^{(t)}$, respectively.  Recall that $\boldsymbol{\mu}^{(t)}$ is only available for sampled grid cells.
\item For unsampled grid cells, generate $\boldsymbol{\mu}^{(t)} \sim \mathcal{N}({\bf X}\boldsymbol{\beta}^{(t)}+{\bf K}\boldsymbol{\alpha}^{(t)},\tau_\epsilon^{(t)})$
\item Sample ${\bf N}^{(t)} \sim \text{Poisson}(\boldsymbol{\mu}^{(t)})$, where ${\bf N}^{(t)}$ is a vector of posterior predictions across the landscape.
\item Compute a posterior prediction for total abundance as $N^{(t)}=\sum_s N_s^{(t)}$.
\end{enumerate}
